|

### 2020-03-11 paper/Manuscript.Rmd
Loek's comments

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"

abstract: |
  The social sciences are amidst a paradigm shift towards open science. In part,
  this transition has been fueled by cases of scientific fraud and increasing
  awareness of questionable research practices. However, Open science is not
  merely a cure (or punishment) for this crisis - it is also an opportunity.
  Technological advances enable researchers to more easily conduct reliable,
  cumulative, collaborative science. Capitalizing on these advances has the
  potential to accelerate scientific progress. This tutorial paper introduces a
  workflow for open reproducible code in science (WORCS). WORCS is a 
  step-by-step procedure that researchers can follow to make an entire project
  Open and Reproducible. It is based on best practices, and can be used either
  in parallel to, or in absence of, top-down requirements by journals. WORCS
  lowers the threshold for adopting an Open Science workflow by providing this
  tutorial, along with an R-package with user-friendly support functions, and an
  Rstudio project template.

@@

<!-- Structure
# Introduction
# Defining Open Science
* TOP guidelines
# Goal of this paper
* Grass-roots
* Open science as opportunity ?move
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction
# Introducing the tools
# Setup - do this only once
# WORCS - steps to follow for each project
# Discussion
* advantages
* limitations
* future directions
-->

@@

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [e.g., @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018]. However, Open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012]. Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012]. Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

This paper introduces WORCS: A Workflow for Open Reproducible Code in Science. WORCS is a lightweight approach to open science and computational reproducibility. It provides a step-by-step procedure that researchers can follow to make an entire research project Open and Reproducible. WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals. The main goal of WORCS is to lower the threshold for adopting an Open Science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.

@@

* Grass-roots
* Open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction

@@

# Defining Open Science

One of the most impactful contributions to Open Science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them. These guidelines describe eight standards for Open Science: 1) Comprehensive citation of literature, data, materials, and methods; sharing 2) data, 3) code required to reproduce analyses, 4) new research materials, and 5) details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results. WORCS defines the goals of Open Science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the topic of a research project, instead of its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations. However, large organizations are slow to change. For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012]. Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support. The formation of "Open Science Communities" at universities around the globe illustrates this movement. This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of Open Science. This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist, facilitates meeting them.

@@

There have been several previous efforts to promote grass-roots adoption of Open Science principles. Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS. For example, there are "signalling solutions"; guidelines to structure and incentivize disclosure about Open Science practices. Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to Open Science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based Transparency Checklist that authors can complete online to generate a report. Such signalling solutions are very easy to adopt, and they address TOP-guidelines 1-7.
Many journals now also offer authors the opportunity to earn "badges" for adhering to Open Science guidelines [@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure authors' disclosures about, and incentivize adherence to, Open Science practices.

A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of Open Science. <!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019. Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5. Peikert and Brandmaier aim to achieve strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and reinstated on different systems, instead of on a private computer. The workflow is excellent, but difficult to implement for non-programmers. WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*. This should be sufficient for most users, but can also be combined with Peikert and Brandmaier if strict computational reproducibility is essential. WORCS thus addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines. Still, it can be combined with existing approaches.

```{r, echo = FALSE, eval = FALSE}
tab1 <- data.frame(
  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
  Scope = c("Disclose adherence to Open Science principles", "Completely reproducible analysis"),
  Difficulty = c("Easy", "Hard"),
  Effectiveness = c("Low", "High"),
  "TOP-guidelines" = c("1-7", "2,3,5")
)
apa_table(tab1)
```

@@

WORCS relies on a few free, open source software solutions that, when combined, facilitate meeting the TOP-guidelines related to citation, sharing, and preregistration. The first tool is *dynamic document generation*. Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off: First, the time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program; second, human error in copy-pasting results is eliminated; third, when revisions require major changes to the analyses, all results, Figures and Tables are automatically updated; and finally, flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. In sum, while writing academic papers in a programming environment might seem counter-intuitive at first, this approach is much more amenable to the needs of academics than most word processing software. It prevents mistakes, and saves time.

@@

The functionality of Git is amplified by services such as GitHub (other services are available). GitHub is best understood as a cloud storage service with social networking functionality. The "cloud storage" aspect of GitHub works as follows: You can "clone" a local Git repository to the GitHub website, as a backup or research archive. The "social network" aspect comes into play when the project is set to "public": This allows other researchers to peruse the repository on the GitHub website to see how the work was done, clone it to their own computer to replicate the original work or apply the methods to their own data, open Issues to ask questions or give feedback on the project, or even send a "Pull request" with changes to the text or code for your consideration. Git and GitHub shine as tools for collaboration, because different people can simultaneously work on different parts of a project, and their changes can be compared and automatically merged on the website. Even on solo projects, working with Git/GitHub has many benefits: Staying organized, being able to start a new study with a clone of an old, similar repository, or splitting off an "experimental branch" to try something new, while retaining the ability to revert to a previous state of the project, or to merge the experimental branch.

@@

Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine. This is quite effective, but very difficult to set up. A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (in my case, hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

@@

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for Open Science. Second, all of the aforementioned tools for an Open Science workflow are already implemented in R. Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Wrappers in R are often available for tools developed in other programming languages. Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the Open Science tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.

Working with R is simplified immensely by using the Rstudio user interface for R. Many of the tools used in WORCS are embedded directly into the visual user interface of Rstudio. Rstudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub. This makes Rstudio a comprehensive solution for Open Science research projects. Another important feature of Rstudio is project management. A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users. The `worcs` R-package installs a new Rstudio project template. When a new project is initialized from this template, much of the bookkeeping required to set up an Open Science project is performed automatically.

Another key consideration is what filetypes to use. A case can be made to use text-based files only (or as much as possible), instead of binary files. Text-based files can be read by machines and humans alike, without requiring special (licensed) software. Binary files must be decoded by software first. Git is designed to keep track of changes to the content of text-based files. Each change reflects a word or line of code added or removed. Binary files such as `.docx` or `.sav`, on the other hand, should not be version controlled. The changes in the encrypted file are uninterpretable. Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon. GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes. For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.

@@

Sometimes, Open Science principles are discussed as a kind of "antidote" to scientific misconduct. Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.

@@

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference. Users can render a manuscript to PDF either with, or without, comprehensive citations.

The procedure for citation in WORCS is as follows:

@@

When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results.

@@

1. Create a new Private repository on github, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
2. In Rstudio, click File > New Project > New directory > WORCS Project Template
    a. Paste the GitHub Repository address in the textbox
    b. Keep the checkbox for `renv` checked if you want to document all dependencies (recommended). The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
    c. Select a preregistration template
    d. Click Open in new session, and Create project
3. Write the preregistration `.Rmd`
4. In the top-right corner of Rstudio, select the Git tab, select the checkboxes next to all files, and click the Commit button. Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
5. Go to the GitHub repository for this project, and tag the Commit as a preregistration
6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
7. Optional: Add study Materials (to which you own the rights) to the repository. It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)

@@

13. Save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
14. Add an Open Science statement to the manuscript, which references the online repository for the project. This reference can be masked for blind review. The Open Science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available. For further guidance, see @aalbersbergMakingScienceTransparent2018. Example:  
    > In the spirit of Open Science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.

@@

Note that some researchers take these steps at an earlier stage. However, for newcomers to Open Science, who are unsure about potential implications for the publication of their work, it might be prudent to perform these actions after acceptance, and before publication:

16. Make the GitHub repository public
17. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
18. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
19. Add an Open Science statement to the manuscript, with a link to the OSF project
20. Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
    + Check [Sherpa Romeo](http://sherpa.ac.uk/romeo/index.php) to be sure that your intended outlet allows the publication of preprints; many journals do, nowadays - and if they do not, it is worth considering other outlets.

@@

In this tutorial paper, we have presented a workflow for open reproducible code in science. The workflow aims to lower the threshold for grass-roots adoption of open science principles. It is accompanied by an R-package with convenience functions, and a WORCS project template for Rstudio. This relatively light-weight workflow meets most of the requirements for Open Science, as detailed in the TOP-guidelines. However, several limitations and issues for future development remain.

>>>>

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Loek Brinkman"
    affiliation   : "2,3"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"
  - id            : "3"
    institution   : "UMC Utrecht"

abstract: |
  The social sciences are amidst a paradigm shift towards open science. In part,
  this transition has been fueled by cases of scientific fraud and increasing
  awareness of questionable research practices and failed replications. However,
  open science is not
  merely a cure (or punishment) for this crisis - it is also an opportunity.
  Technological advances enable researchers to more easily conduct reliable,
  cumulative, collaborative science. Capitalizing on these advances has the
  potential to accelerate scientific progress. This tutorial paper introduces a
  workflow for open reproducible code in science (WORCS). WORCS is a 
  step-by-step procedure that researchers can follow to make an entire project
  Open and Reproducible. It is based on best practices, and can be used either
  in parallel to, or in absence of, top-down requirements by journals. WORCS
  lowers the threshold for adopting an open science workflow by providing this
  tutorial, along with an R-package with user-friendly support functions, and an
  Rstudio project template.

@@

<!-- Structure
# Introduction
# Defining open science
* TOP guidelines
# Goal of this paper
* Grass-roots
* open science as opportunity ?move
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction
# Introducing the tools
# Setup - do this only once
# WORCS - steps to follow for each project
# Discussion
* advantages
* limitations
* future directions
-->

@@

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [e.g., @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018]. However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012]. Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012]. Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

This paper introduces WORCS: A Workflow for Open Reproducible Code in Science. WORCS is a lightweight approach to open science and computational reproducibility. It provides a step-by-step procedure that researchers can follow to make an entire research project Open and Reproducible. WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals. The main goal of WORCS is to lower the threshold for adopting an open science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.

@@

* Grass-roots
* open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction

@@

# Defining open science

<!-- LB: Although Open Science is advocated by many, there is no clear definition of Open Science (ref multiple definitions). An influential operationalisation of Open Science has been de TOP guidlines. -->
One of the most impactful contributions to open science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them. These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results. WORCS defines the goals of open science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the subject of a research project, not to its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations. However, large organizations are slow to change. For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012]. Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support. The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement. This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science. This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.

@@

There have been several previous efforts to promote grass-roots adoption of open science principles. Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS. For example, there are "signalling solutions"; guidelines to structure and incentivize disclosure about open science practices. Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to open science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based Transparency Checklist that authors can complete online to generate a report. Such signalling solutions are very easy to adopt, and they address TOP-guidelines 1-7.
Many journals now also offer authors the opportunity to earn "badges" for adhering to open science guidelines [@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure authors' disclosures about, and incentivize adherence to, open science practices.

A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science. <!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019. Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5. Peikert and Brandmaier aim to achieve strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and reinstated on different systems, instead of on a private computer. The workflow is excellent, but difficult to implement for non-programmers. WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*. This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential. WORCS thus addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines. Still, it can be combined with existing approaches.

```{r, echo = FALSE, eval = FALSE}
tab1 <- data.frame(
  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
  Scope = c("Disclose adherence to open science principles", "Completely reproducible analysis"),
  Difficulty = c("Easy", "Hard"),
  Effectiveness = c("Low", "High"),
  "TOP-guidelines" = c("1-7", "2,3,5")
)
apa_table(tab1)
```

@@

WORCS relies on a few free, open source software solutions that, when combined, facilitate meeting the TOP-guidelines related to citation, sharing, and preregistration. The first tool is *dynamic document generation*. Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off:

* The time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program
* Human error in copy-pasting results is eliminated
* When revisions require major changes to the analyses, all results, Figures and Tables are automatically updated
* Flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. 
* Reproducibility is guaranteed, and can be verified by reviewers and/or readers, because the code is run each time the document is compiled.

In sum, while writing academic papers in a programming environment might seem counter-intuitive at first, this approach is much more amenable to the needs of academics than most word processing software. It prevents mistakes, and saves time.

@@

The functionality of Git is amplified by services such as GitHub (other services are available). GitHub is best understood as a cloud storage service with social networking functionality. The "cloud storage" aspect of GitHub works as follows: You can "clone" a local Git repository to the GitHub website, as a backup or research archive. The "social network" aspect comes into play when the project is set to "public": This allows other researchers to peruse the repository on the GitHub website to see how the work was done, clone it to their own computer to replicate the original work or apply the methods to their own data, open "Issues" to ask questions or give feedback on the project, or even send a "Pull request" with suggested changes to the text or code for your consideration. Git and GitHub shine as tools for collaboration, because different people can simultaneously work on different parts of a project, and their changes can be compared and automatically merged on the website. Even on solo projects, working with Git/GitHub has many benefits: Staying organized, being able to start a new study with a clone of an old, similar repository, or splitting off an "experimental branch" to try something new, while retaining the ability to revert to a previous state of the project, or to merge the experimental branch.

@@

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine. This is quite effective, but very difficult to set up. <!-- Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

@@

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science. Second, all of the aforementioned tools for an open science workflow are already implemented in R. Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Wrappers in R are often available for tools developed in other programming languages. <!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.

Working with R is simplified immensely by using the Rstudio user interface for R. Many of the tools used in WORCS are embedded directly into the visual user interface of Rstudio. Rstudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub. This makes Rstudio a comprehensive solution for open science research projects. Another important feature of Rstudio is project management. A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users. The `worcs` R-package installs a new Rstudio project template. When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.

Another key consideration is what filetypes to use. A case can be made to use text-based files only (or as much as possible), instead of binary files. Text-based files can be read by machines and humans alike, without requiring special (licensed) software. Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`),  must be decoded by software first. Git is designed to keep track of changes to the content of text-based files. Each change reflects a word or line of code added or removed. Binary files, on the other hand, should not be version controlled. The changes in the encrypted file are uninterpretable. Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon. GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes. For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.

@@

<!-- LB:  this section comes at a weird place in the flow of the paper. -->Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct. Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.

@@

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference. Users can render a manuscript to PDF either with, or without, comprehensive citations.

The procedure for citation in WORCS is as follows:<!-- LB: this procedure is secondary to the main workflow. Consider changing the order (see main issue)-->

@@

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results.

@@

1. Create a new Private repository on github, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
2. In Rstudio, click File > New Project > New directory > WORCS Project Template
    a. Paste the GitHub Repository address in the textbox
    b. Keep the checkbox for `renv` checked if you want to document all dependencies (recommended). The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
    c. Select a preregistration template
    d. Click Open in new session, and Create project
3. Write the preregistration `preregistration.Rmd`
4. In the top-right corner of Rstudio, select the Git tab, select the checkboxes next to all files, and click the Commit button. Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
5. Go to the GitHub repository for this project, and tag the Commit as a preregistration
6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
7. Optional: Add study Materials (to which you own the rights) to the repository. It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)

@@

13. Save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
14. Add an open science statement to the manuscript, which references the online repository for the project. This reference can be masked for blind review. The open science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available. For further guidance, see @aalbersbergMakingScienceTransparent2018. Example:  
    > In the spirit of open science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.

@@

Note that some researchers take the following steps at an earlier stage. However, for newcomers to open science, who are unsure about potential implications for the publication of their work, it might be prudent to perform these actions after acceptance, and before publication:

16. Make the GitHub repository public
17. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
18. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
19. Add an open science statement to the manuscript, with a link to the OSF project
20. Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
    + Check [Sherpa Romeo](http://sherpa.ac.uk/romeo/index.php) to be sure that your intended outlet allows the publication of preprints; many journals do, nowadays - and if they do not, it is worth considering other outlets.

@@

In this tutorial paper, we have presented a workflow for open reproducible code in science. The workflow aims to lower the threshold for grass-roots adoption of open science principles. It is accompanied by an R-package with convenience functions, and a WORCS project template for Rstudio. This relatively light-weight workflow meets most of the requirements for open science, as detailed in the TOP-guidelines. However, several limitations and issues for future development remain.
